---
title: "p8105_hw3_sjk2254"
author: "Sharon Kulali"
output: github_document
---

## SetUp

```{r, message = FALSE}
# loading the needed libraries

library(tidyverse)
library(p8105.datasets)
library(patchwork)
```

```{r}
# setting visuals for plots

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

```{r}
# loading the dataset

data("instacart")

# looking at the dataset

insta_df = instacart |> 
  as_tibble()
```

Instacart is an online grocery shopping service that allows individuals to shop online from local stores. The platform offers same-day delivery with purchases being able to be delivered within 2 hours. The `insta_df` dataset contains the online grocery orders from Instacart users in 2017. There are `r ncol(insta_df)` variables with `r nrow(insta_df)` observations. More specifically, the dataset has information about identifiers for the order (`order_id`), product (`product_id`), customer (`user_id`), aisle (`aisle_id`), and department (`department_id`). Additionally, there is information about the specific day (`order_dow`) and hour (`order_hour_of_day`) that the orders were placed.

```{r}
# examining the aisles

insta_df |> 
  group_by(aisle_id, aisle) |> 
  count() |> 
  arrange(desc(n))
```

The table above shows the the aisles arranged in decreasing order by most items. Overall, there are 134 different aisles with fresh vegetables and fruits being the most ordered items.

```{r}
# plotting the number of items ordered per aisle

insta_df |> 
  group_by(aisle) |> 
  count() |> 
  filter(n > 10000) |>
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust=1)) +
  ggtitle("Number of Items Ordered in Each Aisle") +
  theme(plot.title = element_text(hjust = 0.5))
```

The plot above shows the number of items ordered in each aisle for aisles with more than 10,000 items ordered. You can see that fresh vegetables and fruits are the most ordered items because they have the highest points.

```{r, message = FALSE}
insta_df |> 
  filter(aisle %in% c("baking ingredients","dog food care","packaged vegetables fruits")) |>  
  group_by(aisle) |> 
  count(product_name) |> 
  arrange(desc(n)) |> 
  top_n(n = 3) |> 
  knitr::kable()
```

The table above shows three most popular items in each of the aisles baking ingredients, dog food care, and packaged vegetables fruits. You can see that packaged vegetables fruits has the most orders out of the three aisles.

```{r, message = FALSE}
insta_df |> 
  filter(product_name %in% c("Pink Lady Apples","Coffee Ice Cream")) |> 
  group_by(product_name, order_dow) |> 
  summarize(mean_hr = mean(order_hour_of_day)) |> 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hr) |> 
  knitr::kable()
```

The table above shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. You can see that on average, Coffee Ice Cream is ordered more than Pink Lady Apples.